---
title: "Data preparation"
author: "Ich"
date: "3/9/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r knitr-setup}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "70%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp =  0.4,  #0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)

```



# Load packages


```{r}
library(sjmisc)
library(viridis)
library(pradadata)  # elec data
library(tidyverse)
library(knitr)
library(mosaic)
library(moments)  # skewness, kurtosis

library(rethinking)  # Bayes modeling

```


# Load data


```{r}
d <- read_csv("objects/d_joined_values.csv")
```


# Rename variables

Value type variable names:

```{r}
value_types <- c("Genieser", "Harmoniesuchende", "Hedonisten", 
                 "Selbstbestimmte", "Werschätzenden", 
                 "type_unknown", "Konformisten",
                 "Verantwortungsverweiterer")
```

Rename all seven types:

```{r}
d_renamed <- d %>% 
  rename(enjoyer = Genieser,
         harmony_seeker = Harmoniesuchende,
         hedonist = Hedonisten,
         self_determined = Selbstbestimmte,
         appreciater = Werschätzenden,
         conformist = Konformisten,
         responsibility_denier = Verantwortungsverweiterer)
```


Here's the list as a string vector:

```{r}
consumer_types <- c("enjoyer", "harmony_seeker",
                    "hedonist", "self_determined",
                    "appreciater", "type_unknown",
                    "conformist", "responsibility_denier")
```





# z-standardize


```{r d-short-z}
d_renamed %>% 
  sjmisc::std() %>%  
  select(-c(state_z, area_nr_z, area_name_z, 
            state_id_z, east_z, east_num_z, 
            paste0(consumer_types, "_z"))) -> d_short_z

names(d_short_z)
```



# Check distribution of afd_prop (outcome)

## afd_prop raw

```{r results = "hold"}
gf_dens(~afd_prop, data = d_short_z)
gf_qqline(~afd_prop, data = d_short_z) %>% 
  gf_qq(~afd_prop)
```

```{r}
skewness(d_short_z$afd_prop)
kurtosis(d_short_z$afd_prop)
```


That's not normally distributed. Let's try to normalize it:


## afd_prop log

```{r results = "hold"}
gf_dens(~log(afd_prop), data = d_short_z)
gf_qqline(~log(afd_prop), data = d_short_z) %>% 
  gf_qq(~log(afd_prop), data = d_short_z)
```

```{r}
skewness(log(d_short_z$afd_prop))
kurtosis(log(d_short_z$afd_prop))
```

Skewness is good; kurtosis is good, since the kurtosis of the normal distribution is 3!

Let's add this variable to thes datafrae:

```{r}
d_short_z <- d_short_z %>% 
  mutate(afd_prop_log = log(afd_prop))
```



## afd_prop sqrt

```{r results = "hold"}
gf_dens(~sqrt(afd_prop), data = d_short_z)
gf_qqline(~sqrt(afd_prop), data = d_short_z) %>% 
  gf_qq(~sqrt(afd_prop), data = d_short_z)
```


## afd_prop $x^-1$

```{r}
gf_dens(~(afd_prop^-1), data = d_short_z)
gf_qqline(~(afd_prop^-1), data = d_short_z) %>% 
  gf_qq(~(afd_prop^-1))

```


## afd_prop $x^-2$

```{r}
gf_dens(~(afd_prop^-2), data = d_short_z)
gf_qqline(~(afd_prop^-2), data = d_short_z) %>% 
  gf_qq(~(afd_prop^-2))

```



# Save data


```{r}
write_csv(d_short_z,
          path = "objects/d_prepared.csv")
```


