---
title: "Join zip data to values data"
date: "19 11 2018"
output: html_document
params:
  date: !r Sys.Date()
  verbose: TRUE
editor_options: 
  chunk_output_type: console
---



```{r knitr-setup}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  out.width = "70%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp =  0.4,  #0.618,  # 1 / phi
  fig.show = "hold",
  size = "tiny"
)

```



# Load packages


```{r}
library(sjmisc)
library(viridis)
library(pradadata)  # elec data
library(tidyverse)
```




# Load data



## Zip data
Load zip data (PLZ):


```{r load-zip-data}
zip <- readxl::read_excel("data_raw/Wahlkreise_PLZ.xlsx")
```


This dataset was accessed form [DeStatis](https://www.destatis.de/DE/ZahlenFakten/LaenderRegionen/Regionales/Gemeindeverzeichnis/Administrativ/Archiv/GVAuszugQ/BTW20174Q2016.html)


Copyright Notice:

 Daten (im Auftrag der Herausgebergemeinschaft Statistische Ämter des Bundes und der Länder)
      Statistisches Bundesamt, Wiesbaden 2017
           Vervielfältigung und Verbreitung, auch auszugsweise, mit Quellenangabe gestattet.
           
          
          
          
## Values data

Load values data:


```{r load-values-data}
load("data_raw/Choughs2017.Rdata")
values <- Choughs2017
```



## Election data

These data are provided by the Bundeswahlleiter 2017.


```{r data-elec}
data("elec_results")
```


In order to combine socioeconomic data with the election results, we can make use of data from the same source as above. Again accessible via the same R pacakge:

```{r data-socec}
data("socec")
```

Note that a code book is available for these data:

```{r data-socec-dict}
data("socec_dict")

```



These data will be used as predictors for modeling the election results.

Third, we will make use of geo data in order to geoplot the modeling results. The Bundeswahlleiter provides such data (again via `pradadata`):

```{r data-shp}
data("wahlkreise_shp")

```



Note: Data objects can also be downloaded [from this source](https://osf.io/2yhr9/).



# Data preparation


Now let's merge the data frames. There will also be some janitor work such as renaming columns etc.


First, change the names of the `socec` data to a common format:

```{r soceec-rename}
socec_renamed <- socec %>%
  rename(state = V01,
         area_nr = V02,
         area_name = V03,
         total_n = V06,
         germans_n = V07,
         for_prop = V08,
         pop_move_prop = V11,
         pop_migr_background_prop = V19,
         income = V26,
         unemp_prop = V47) 

```


Compute some more columns:

```{r socec-new-cols}
socec2 <- socec_renamed %>% 
   mutate(foreigner_n = total_n - germans_n,
         pop_move_n = pop_move_prop * total_n,
         unemp_n = unemp_prop * total_n / 100,
         pop_migr_background_n = pop_migr_background_prop * total_n / 100) %>% 
  drop_na()
```



Same thing with the election data, here we only need the criterion (AfD success) and the ID variables for merging:

```{r elec-results-new-cols}
elec_results2 <- elec_results %>%
  rename(afd_votes = AfD_3,
         area_nr = district_nr,
         area_name = district_name,
         votes_total = votes_valid_3) %>% 
   mutate(afd_prop = afd_votes/votes_total)    # valid votes only, and of the present Zweitstimme
```

Note that we are focusing on the Zweitstimme of the present election (hence the `3` in `votes_valid_3` and in `AfD_3`).

# Merge


```{r merge}
socec2 %>%
  left_join(elec_results2, by = "area_name") %>% 
  left_join(wahlkreise_shp, by = c("area_name" = "WKR_NAME")) -> d_all_with_na
```


# After-merge preparations

<!-- Now let's remove missings: -->

<!-- ```{r drop-na} -->
<!-- d_all <- drop_na(d_all_with_na) -->
<!-- ``` -->

Add variable for East (1) vs. West (0):


```{r east-west-dummy}
d_all_with_na <- d_all_with_na %>% 
  mutate(east = case_when(
    state %in% c("Mecklenburg-Vorpommern", "Brandenburg", "Berlin", "Sachsen-Anhalt", "Sachsen", "Thüringen") ~ "yes",
    TRUE ~ "no"
    ) 
  )

d_all_with_na$east_num <- ifelse(d_all_with_na$east == "yes", 1, 0)
```



# Main data frame: d_short and d_short_X

We will also provide a version without the geo data, and in pure (old school) `data frame` form (ie., not as tibble)_

```{r d-short-with-nas}
d_all_with_na %>%
  rename(area_nr = area_nr.x) %>% 
  select(state,
         area_nr,
         area_name,
         total_n,
         germans_n,
         foreigner_n,
         for_prop,
         pop_move_n,
         pop_migr_background_n,
         income ,
         unemp_n,
         unemp_prop,
         votes_total,
         afd_votes,
         afd_prop,
         state,
         east,
         east_num) -> d_short_with_nas

if (params$verbose == TRUE) names(d_short_with_nas)
```


Remove NAs:

```{r nona}
d_short_with_nas %>% 
  drop_na() -> d_short_nona
```


Add state id:

```{r staste-id}
d_short_nona$state_id <- coerce_index(d_short_nona$state)
```





Multiply by 1000 to get the real numbers so that a count model gets the "right" data

```{r mult-by-1000}
d_short_nona %>%
  mutate_at(vars(total_n, germans_n, foreigner_n, pop_move_n,
                    pop_migr_background_n, unemp_n), funs(. * 1000)
  ) -> d_short_nona_1000

if (params$verbose == TRUE) glimpse(d_short_nona_1000)

```




# Join election data with zip data



```{r join-elec-with-zip}
d_zip <- d_short_nona %>% 
  left_join(zip, by = c("area_name" = "Wahlkreis_Name"))
```


Note that a left join was performed.

# Join election data with values data

```{r join-elec-with-values}

d_joined <- d_zip %>% 
  left_join(values, by = c("PLZ" = "PLZ"))
```



# Check

Number of rows in the final dataset should approximate the row count in the values data set.

```{r}
nrow(values)/nrow(d_joined)
```


## Number of missings per column


```{r}
d_joined %>% 
  map(~ sum(is.na(.)))
```


## Number of missings per row

```{r}
d_joined %>% 
  mutate(NA_row = rowSums(is.na(.))) %>% 
  ggplot(aes(x = NA_row)) +
  geom_histogram()
```


# Summarise zu PLZ level


```{r}

```

